---
title: "Company Bankruptcy - Regression Analysis"
fontsize: 10pt
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
---

## Loading Dataset

```{r message=FALSE}
library(DT)
df<-read.csv("C:/Users/angel/Downloads/data.csv")
X<-df[-c(1)]
y<-df[1]
datatable(head(df,50))
```
Y - Bankrupt?: Class label  
X1 - ROA(C) before interest and depreciation before interest: Return On Total Assets(C)  
X2 - ROA(A) before interest and % after tax: Return On Total Assets(A)  
X3 - ROA(B) before interest and depreciation after tax: Return On Total Assets(B)  
X4 - Operating Gross Margin: Gross Profit/Net Sales  
X5 - Realized Sales Gross Margin: Realized Gross Profit/Net Sales  
X6 - Operating Profit Rate: Operating Income/Net Sales  
X7 - Pre-tax net Interest Rate: Pre-Tax Income/Net Sales  
X8 - After-tax net Interest Rate: Net Income/Net Sales  
X9 - Non-industry income and expenditure/revenue: Net Non-operating Income Ratio  
X10 - Continuous interest rate (after tax): Net Income-Exclude Disposal Gain or Loss/Net Sales  
X11 - Operating Expense Rate: Operating Expenses/Net Sales  
X12 - Research and development expense rate: (Research and Development Expenses)/Net Sales  
X13 - Cash flow rate: Cash Flow from Operating/Current Liabilities  
X14 - Interest-bearing debt interest rate: Interest-bearing Debt/Equity  
X15 - Tax rate (A): Effective Tax Rate  
X16 - Net Value Per Share (B): Book Value Per Share(B)  
X17 - Net Value Per Share (A): Book Value Per Share(A)  
X18 - Net Value Per Share (C): Book Value Per Share(C)  
X19 - Persistent EPS in the Last Four Seasons: EPS-Net Income  
X20 - Cash Flow Per Share  
X21 - Revenue Per Share (Yuan ¥): Sales Per Share  
X22 - Operating Profit Per Share (Yuan ¥): Operating Income Per Share  
X23 - Per Share Net profit before tax (Yuan ¥): Pretax Income Per Share  
X24 - Realized Sales Gross Profit Growth Rate  
X25 - Operating Profit Growth Rate: Operating Income Growth  
X26 - After-tax Net Profit Growth Rate: Net Income Growth  
X27 - Regular Net Profit Growth Rate: Continuing Operating Income after Tax Growth  
X28 - Continuous Net Profit Growth Rate: Net Income-Excluding Disposal Gain or Loss Growth  
X29 - Total Asset Growth Rate: Total Asset Growth  
X30 - Net Value Growth Rate: Total Equity Growth  
X31 - Total Asset Return Growth Rate Ratio: Return on Total Asset Growth  
X32 - Cash Reinvestment %: Cash Reinvestment Ratio  
X33 - Current Ratio  
X34 - Quick Ratio: Acid Test  
X35 - Interest Expense Ratio: Interest Expenses/Total Revenue  
X36 - Total debt/Total net worth: Total Liability/Equity Ratio  
X37 - Debt ratio %: Liability/Total Assets  
X38 - Net worth/Assets: Equity/Total Assets  
X39 - Long-term fund suitability ratio (A): (Long-term Liability+Equity)/Fixed Assets  
X40 - Borrowing dependency: Cost of Interest-bearing Debt  
X41 - Contingent liabilities/Net worth: Contingent Liability/Equity  
X42 - Operating profit/Paid-in capital: Operating Income/Capital  
X43 - Net profit before tax/Paid-in capital: Pretax Income/Capital  
X44 - Inventory and accounts receivable/Net value: (Inventory+Accounts Receivables)/Equity  
X45 - Total Asset Turnover  
X46 - Accounts Receivable Turnover  
X47 - Average Collection Days: Days Receivable Outstanding  
X48 - Inventory Turnover Rate (times)  
X49 - Fixed Assets Turnover Frequency  
X50 - Net Worth Turnover Rate (times): Equity Turnover  
X51 - Revenue per person: Sales Per Employee  
X52 - Operating profit per person: Operation Income Per Employee  
X53 - Allocation rate per person: Fixed Assets Per Employee  
X54 - Working Capital to Total Assets  
X55 - Quick Assets/Total Assets  
X56 - Current Assets/Total Assets  
X57 - Cash/Total Assets  
X58 - Quick Assets/Current Liability  
X59 - Cash/Current Liability  
X60 - Current Liability to Assets  
X61 - Operating Funds to Liability  
X62 - Inventory/Working Capital  
X63 - Inventory/Current Liability  
X64 - Current Liabilities/Liability  
X65 - Working Capital/Equity  
X66 - Current Liabilities/Equity  
X67 - Long-term Liability to Current Assets  
X68 - Retained Earnings to Total Assets  
X69 - Total income/Total expense  
X70 - Total expense/Assets  
X71 - Current Asset Turnover Rate: Current Assets to Sales  
X72 - Quick Asset Turnover Rate: Quick Assets to Sales  
X73 - Working capitcal Turnover Rate: Working Capital to Sales  
X74 - Cash Turnover Rate: Cash to Sales  
X75 - Cash Flow to Sales  
X76 - Fixed Assets to Assets  
X77 - Current Liability to Liability  
X78 - Current Liability to Equity  
X79 - Equity to Long-term Liability  
X80 - Cash Flow to Total Assets  
X81 - Cash Flow to Liability  
X82 - CFO to Assets  
X83 - Cash Flow to Equity  
X84 - Current Liability to Current Assets  
X85 - Liability-Assets Flag: 1 if Total Liability exceeds Total Assets, 0 otherwise  
X86 - Net Income to Total Assets  
X87 - Total assets to GNP price  
X88 - No-credit Interval  
X89 - Gross Profit to Sales  
X90 - Net Income to Stockholder's Equity  
X91 - Liability to Equity  
X92 - Degree of Financial Leverage (DFL)  
X93 - Interest Coverage Ratio (Interest expense to EBIT)  
X94 - Net Income Flag: 1 if Net Income is Negative for the last two years, 0 otherwise  
X95 - Equity to Liability  

## Summary for each variable
```{r}
library(skimr)
skim(df)
```

## Correlation among variables
In order to fit linear regression, we need some appropriate correlation among variables 
```{r}
corr_matrix <- round(cor(df[-c(1)]), 2) #drop the target column
which(abs(corr_matrix) > 0.95 & corr_matrix < 1, arr.ind=TRUE)
```
## Linear Regression with one predictor

```{r}
plot(X$Net.worth.Assets, X$Current.Liability.to.Assets)
plot(X$Net.worth.Assets, X$Equity.to.Liability)
plot(X$Total.Asset.Turnover, X$Cash.Total.Assets)
plot(X$Persistent.EPS.in.the.Last.Four.Seasons, X$Per.Share.Net.profit.before.tax..Yuan.Â..) #EPS- Earnings per share
```
The second model is non-linear model called *intrinsically linear*.
It is linear cause it can be transformed to straight line by a logarithmic transformation.
```{r}
df.copy<- df[df$Equity.to.Liability > 0.1,]
X.copy <- df.copy[-c(1)]
exponential.model <- lm(log(X.copy$Equity.to.Liability) ~ X.copy$Net.worth.Assets)
summary(exponential.model)
summary(lm(X.copy$Equity.to.Liability~X.copy$Net.worth.Assets))
plot(exponential.model)

#The output was error cause there are x and y values that are not in domain of log-function.
# Reminder: remove the outlier where Y=0.0 that is showed in second scatter plot.
```
### Interpreting the intercept and the slope in a regression model
```{r}
fit<-lm(X$Persistent.EPS.in.the.Last.Four.Seasons~X$Per.Share.Net.profit.before.tax..Yuan.Â..)
summary(fit)
```
There is no special interpreting on intercept. There will be if value of X is0, then intercept is expected value of Y.

If the ratio of the variable 'Earnings Per Share-Net Income' increases by 1 unit value, then the variable 'Pretax Income Per Share' will increase by approximately 0.96. This is interpretation of the slope  

### Residuals
```{r}
y<- X$Per.Share.Net.profit.before.tax..Yuan.Â..
x<- X$Persistent.EPS.in.the.Last.Four.Seasons
fit2<-lm(y~x)
# With implemented method
e<- resid(fit2)
#Now we're going to obtain the residuals manually, firs we need to get Ys by predicting
yhat<-predict(fit2)
#Residuals are defined by y-yhat
#Let's check by comparing this with R's build in resid function
max(abs(e-(y-yhat)))
#Now, with hard coding the calculation of yhat
max(abs(e-(y- coef(fit2)[1]-coef(fit2)[2]*x)))
```
### Residuals versus X
Using this plot, we can see the instances of poor model fit
```{r}
plot(x, resid(lm(x~y)))
plot(X.copy$Net.worth.Assets, X.copy$Equity.to.Liability) #nonlinear model
abline(lm(X.copy$Equity.to.Liability ~ X.copy$Net.worth.Assets ), col='red')
```
Here's what happens when we focus on the residuals  

```{r}
y<- X.copy$Net.worth.Assets
x<- X.copy$Equity.to.Liability
library(ggplot2)
ggplot(X.copy, aes(Equity.to.Liability, resid(lm(y~x)))) + 
    geom_point(size = 3, col = "firebrick") + 
    geom_hline(yintercept=0, color = "black") +
    labs(x = "Equity.to.Liability", y = "Residuals") + 
    theme_classic()
```  
Estimating residual variation
```{r}
fit<-lm(y~x); n<-length(y)
#Comparing the estimate from lm comparing and  directly calculating from the residuals
summary(fit)$sigma == sqrt(sum(resid(fit)^2)/(n-2))
```
### Summarizing variation
```{r}
y<- X$Per.Share.Net.profit.before.tax..Yuan.Â..
x<- X$Persistent.EPS.in.the.Last.Four.Seasons
fit<-lm(y~x)
regress.var = sum((predict(fit)-mean(y))^2)
residual.var = sum((y-predict(fit))^2)
total.var= regress.var+residual.var
total.var
#To check, derive from equation for R squared R^2 = regress.var / total.var6
regress.var/summary(fit)$r.squared
```
## Regression inference
```{r}
y<-X$Current.Liability.to.Assets; x<-X$Net.worth.Assets; n<-length(X$Net.worth.Assets)
beta1<- cor(y,x)*sd(y)/sd(x)
beta0<- mean(y)-beta1*mean(x)
e <- y - beta0 - beta1*x
sigma<- sqrt(sum(e^2)/(n-2))
ssx<- sum((x-mean(x))^2) #sum of squared deviations from the mean of X

#Calculate standard errors for the coefficients and t-statistic
#Null hypotheses are Ho : Bj=0
seBeta0<- (1 / n + mean(x) ^ 2 / ssx) ^ 0.5 * sigma
seBeta1<- sigma/sqrt(ssx)
tBeta0 <- beta0 / seBeta0
tBeta1 <- beta1 / seBeta1

#p-values
pBeta0 <- 2 * pt(abs(tBeta0), df = n-2, lower.tail = FALSE)
pBeta1 <- 2 * pt(abs(tBeta1), df = n-2, lower.tail = FALSE)
coefTable<-rbind(c(beta0,seBeta0,tBeta0,pBeta0), c(beta1, seBeta1, tBeta1, pBeta1))
colnames(coefTable)<- c("Estimates", "Std. Error", "t value", "P(>|t|)")
rownames(coefTable)<- c("(Intercept)","x")
coefTable

#built-in methods
fit<-lm(y~x)
summary(fit)$coefficients
```
### Getting a confidence interval
Confidence interval for the intercept
```{r}
sumCoef<- summary(fit)$coefficients
sumCoef[1,1]+c(-1,1)*qt(0.95, df=fit$df)*sumCoef[1,2]
```
Now, let's do the slope:
```{r}
sumCoef[2,1]+c(-1,1)*qt(0.95, df=fit$df)*sumCoef[2,2]
```
We would interpret this as: **"with 95% confidence, we estimate that if Net Worth Assets ratio increase by 1, the Current Liability to Assets ratio will decrease by approximately -0.79 to -0.77"** 
```{r}
library(ggplot2)
ggplot(df[1:200,], aes(x=Net.worth.Assets, y=Current.Liability.to.Assets)) + 
  geom_point(color='#2980B9', size = 4) + 
  geom_smooth(method=lm, color='#2C3E50')
#The gray bands around the line represent the standard error of the regression line
```
## Multivariable regression analysis
```{r}
#Targer->compute beta1
x1<-X$Revenue.per.person
x2<-X$Working.Capital.Equity
x3<-X$Fixed.Assets.to.Assets
y<-X$Research.and.development.expense.rate
#Get the residuals having removed X2 and X3 from X1 and Y
ey<-resid(lm(y ~ x2 + x3))
ex1<-resid(lm(x1 ~ x2 + x3))
#Fit regression through the origin with the residuals
sum(ey * ex1)/sum(ex1 ^ 2)
#Check with built-in method lm
coef(lm(ey ~ ex1 - 1))
#Fit the full linear model to show that it agrees
coef(lm(y~x1+x2+x3))
```
```{r}
#Examples with more features
data<- df[,c("Working.Capital.to.Total.Assets","Quick.Assets.Total.Assets","Current.Assets.Total.Assets","Cash.Flow.to.Sales","Working.capitcal.Turnover.Rate","Debt.ratio..","Net.worth.Assets")]
summary(lm(Working.Capital.to.Total.Assets ~ ., data=data))
summary(lm(Quick.Assets.Total.Assets ~ . - Net.worth.Assets, data=data))
```
## Dummy variables
```{r}
library(dplyr)
data = mutate(data, RevenuePerEmployee = 1 * (X$Revenue.per.person > 0.018))
library(ggplot2)
g = ggplot(data[1900:2200,], aes(x=Working.Capital.to.Total.Assets, y=Current.Assets.Total.Assets, color=factor(RevenuePerEmployee)))
g = g + geom_point(size=3, colour = "black") + geom_point(size=2)
g
```
Thus, -0.14 is the estimated change in the intercept of the linear relationship between Working Captial to Total Assets and Current.Assets.Total.Assets when Revenue per person is high.
The interaction term, 0.287, is the estimate change in the slope. The estimated intercept is -1.48 when Revenue per person is low, while the estimated intercept is -1.48 + 0.287. The estimated slope is 2.39 if Revenue per person is low, else 2.39 + 0.287. Since, all of the main effects are change in the intercepts from the reference level, while all of the interaction terms are changes in slope.

```{r}
summary(lm(Current.Assets.Total.Assets ~ Working.Capital.to.Total.Assets * factor(RevenuePerEmployee),data=data))$coef
#The interaction term is the estimate change in the slope.
```
## Residuals, variation, diagnostic
```{r}
plot(lm(Working.Capital.to.Total.Assets ~ ., data=data))
```
### Outliers, Leverage, Influence
```{r}
fit<-lm(Working.Capital.to.Total.Assets ~ Net.worth.Assets, data=data)
round(dfbetas(fit)[1:10,2],3) #change in individual coefficients when i-th point is deleted in fitting model
round(hatvalues(fit)[1:10],3)# change in the predicted response when i-th point is deleted in fitting the model
```
## Variance Inflation Factor
```{r}
library(car)
fit<-lm(Working.Capital.to.Total.Assets ~ . - Net.worth.Assets, data=data)
vif(fit)
# measure of the amount of multicollinearity in a set of multiple regression variables
```
## Principal Component Analysis
```{r}
companies.pca<-prcomp(df[, c(30:35)],center = TRUE, scale. = TRUE)
summary(companies.pca)
```
```{r}
library(devtools)
install_github("vqv/ggbiplot")
library(ggbiplot)
ggbiplot(companies.pca) 
```

## Logistic Regression
```{r}
glm.fit<-glm(df$Bankrupt~companies.pca$x)
summary(glm.fit)
```

The residual deviance shows how well the response is predicted by the model when the predictors are included.  

The null deviance shows how well the response is predicted by the model with nothing but an intercept.  
This increase in deviance is evidence of a significant lack of fit.

We can also use the residual deviance to test whether the null hypothesis is true (i.e. Logistic regression model provides an adequate fit for the data). This is possible because the deviance is given by the chi-squared value at a certain degrees of freedom.

## Poisson regression
```{r}
covid19<-read.csv("C:/Users/angel/Downloads/data (1).csv")
summary(fit<-glm(abs(deaths) ~ cases, family="poisson", data=covid19[covid19$geoId=="AT",]))
plot(fit)
#y- deaths , x - cases
```
